---
title: "A guide to using 'ptspotter' v1.0.1"
author: "Rich Leyshon"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = tempdir())
```

## What is 'ptspotter'?

It's a package with helper functions, designed to work well with 'ProjectTemplate'. However, the functions' utility are not limited to use with 'ProjectTemplate'. 'ProjectTemplate' is an excellent package that can be used to organise and efficiently set up more complex workflows. I highly recommend this package. Please check it out on [CRAN](https://CRAN.R-project.org/package=ProjectTemplate) and also the super useful [guide to getting started](http://projecttemplate.net/getting_started.html).

***

## Why 'ptspotter'?

It's a bad joke. Reflecting on the CamelCase syntax for 'ProjectTemplate' and that PT tends to mean Personal Trainer. Who would help a Personal Trainer, maybe a spotter?

***

## 'ptspotter' Functions Overview.

You may find you are spending a disproportionate amount of time manually renumbering R scripts when features are added or removed from your written procedures. `seq_file_ops()` Can be used to quickly create sequentially numbered scripts and `adj_file_nos()` can be used to adjust their numbering.

`memory_report()` and `wrap_up()` are thin wrapper functions that capture some common functions that can be used for debugging procedures. You can use `memory_report()` to monitor memory allocation at select points within your written procedure. `wrap_up()` is used to `stop()` a procedure at any point but has the advantage of using the excellent ['this.path'](https://CRAN.R-project.org/package=this.path) package to help you pinpoint (and remember) which R script you called `wrap_up()` from.

When logging a procedure with ['log4r'](https://cran.r-project.org/web/packages/log4r/index.html), there are a few setup steps to take in getting everything up and running. 'ProjectTemplate' does a great job of making this pain free. However I don't always use it, expecially if a project is smaller and doesn't merit the more formalised approach that 'ProjectTemplate' adheres to. `log_file_ops()` is used to setup the directory and logfile prerequisites. `log_enable()` is used to setup the 'log4r' logger and file appender requirements. These 2 functions allow you to quickly get logging any R workflow.

***

## How to get 'ptspotter'.

Installing the stable version from CRAN.

```{r, eval=FALSE}
install.packages("ptspotter")
```

Loading the development version from GitHub.

```{r, eval=FALSE}
devtools::install_github("r-leyshon/ptspotter")
```

## Using 'ptspotter' Functions.

Load the package.

```{r, message=FALSE}
library(ptspotter)
```


### File Operations.

To quickly produce the required number of sequential scripts within a directory of your choice.

```{r}
seq_file_ops(n = 10, target_dir = "munge")
list.files("munge")
```

Note that `seq_file_ops()` creates the directory specified with `target_dir` if it does not already exist.

At some point in the development process, features may be added to the procedure. This may leave you needing to renumber the scripts. Instead of manually renumbering these scripts, `adj_file_nos()` can do the work for you. In the following example, I need to add 2 more scripts halfway through the sequence.

```{r}

adj_file_nos(target = 5, directory = "munge", step = 2)

list.files("munge")
```

You can see I now have a gap in my sequence. To fill it, I can rerun the `seq_file_ops()` command run earlier. **But wait, won't that write over my existing scripts?** I hear you ask. Not if you keep the `force = FALSE`, the default value.

```{r}
seq_file_ops(n = 10, target_dir = "munge", force = FALSE)
```

As you can see, `seq_file_ops()` tells you that the pre-existing files in your specified sequence have not been overwritten. 

Now let's imagine those new features were really not such a great idea after all. Let's remove the 2 files we added.

```{r}
unwanted <- paste("munge", c("05-.R", "06-.R"), sep = "//")
file.remove(unwanted)

list.files("munge")
```
We have that gap in the sequence again. To correct this, we can use `adj_file_nos()` once more, this time specifying the argument `action = "down"`

```{r}
adj_file_nos(target = 5, directory = "munge", action = "down", step = 2)

list.files("munge")
```


```{r, include=FALSE}
unlink("munge", recursive = TRUE)
```





